---
title: "Weather Prediction"
author: "Cody Overholt"
date: "9/18/2021"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(pander)

weatherData <- read.csv("C:/Users/Cody/Documents/College/10th semester (fall 2021)/Math 425 (linear Regression)/Statistics-Notebook-master/Statistics-Notebook-master/Analyses/Linear Regression/WeatherData.csv")
targetData <- read.csv('C:/Users/Cody/Documents/College/10th semester (fall 2021)/Math 425 (linear Regression)/Statistics-Notebook-master/Statistics-Notebook-master/Analyses/Linear Regression/target weather.csv')



#averages the daily high temperatures for the past 3 days
aggWeathDat <- weatherData %>% 
  select(c(year,High)) %>% 
  group_by(year) %>% 
  summarise(mean = mean(High))
  
#joins the datasets
completeDat <- aggWeathDat %>% 
  left_join(targetData, by = "year") %>% 
  filter(year<2021)

weatherLM <- lm(high~mean,data=completeDat)
weatherLM

summary(weatherLM)

predVal <- predict(weatherLM, data.frame(mean = 77))
predVal

#View Datasets
View(weatherData)
View(targetData)
View(aggWeathDat)
View(completeDat)
```

## Prediciton

The expected daily high temperature for 9/20/2021 is 71.44 degrees Fahrenheit given the regression model below.


```{r}
#graph of the reggression model
completeDat %>% 
  ggplot(aes(x=mean,y=high))+
  geom_point(alpha =1,color="skyblue", fill = "black")+
  geom_smooth(method="lm", color="black", size = .5)+
  geom_point(aes(x=77,y=predVal, size = 1.25), color="firebrick",show.legend=FALSE)+
  geom_text(x=77,y=85, label="Predicted Temperature for 9/20/2021 \n 77 Degrees", color="black")+
  geom_segment(x=77,y=predVal+1,xend=77,yend=83)+
  labs(title="High Temperature Prediction for 9/20", x="Average High Temperature for 9/16 - 9/18", y="High Temperature for 9/20")+
  theme_bw()

```

Given the lack of data, the prediction made by this model should not be trusted for accurate results. There is a large amount of variance that decreases its overall accuracy and usefulness.



# Technical Analysis

## Hypothesis and Background

The data for this study came from weather data collected from https://www.timeanddate.com/weather/@5605242/historic?month=1&year=2019 for Rexburg Idaho. The data used ranged from 2011 to 2021 in order to predict the daily high temperature for September 20th, 2021. The following model was used to analyze the dataset:

$$
  \underbrace{Y_i}_\text{9/20 High temp} = \beta_0 + \overbrace{\beta_1}^\text{slope} \underbrace{X_i}_\text{Average temp9/16-9/18} + \epsilon_i \quad \text{where} \ \epsilon_i \sim N(0, \sigma^2)
$$

## Results

The results of the model are as follows:

```{r}
summary(weatherLM) %>% pander()

```

Looking at the P-values of the slope and intercepts, the model did not produce any statistically significant results. Additionally, none of the assumptions were met. Residuals do not show constant variance and many of the values fall outside of the normal bounds.

```{r}

par(mfrow=c(1,3))
plot(weatherLM, which=1:2)





#mtext(side=3,text="Q-Q Plot of Residuals")
plot(weatherLM$residuals)
#mtext(side=3, text="Residuals vs. Order")
```

## Conclusion

Given the lack of statistical significance and failure to meet any of the assumptions, the model presented here is inadequate to be used in any legitimate capacity. One potential avenue to explore to improve the model is to increase the size of the dataset. This could be done by including more years in the data or including more timeframes for prediction.
