---
title: "Recovering a True Model"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
p1Data <- read.csv("p1Data-1.csv", header=TRUE)
p2Data <- read.csv("p2Data-1.csv", header=TRUE)
p3Data <- read.csv("p3Data-1.csv", header=TRUE)

```

```{r}
#shows patterns between variables
#Ignore bottom and left
pairs(p1Data)


#shows potential pattern lines
pairs(p1Data,panel = panel.smooth)
#colors by X2 variable
pairs(p1Data,panel = panel.smooth, col=as.factor(p1Data$X2))

#obvious 2 lines model, but looking as if we didnt know
mylm1 <- lm(Y~X4,data=p1Data)
summary(mylm1)

#the dot iterates along each column in dataset
plot(mylm1$residuals~.,p1Data)

#we saw that X2 had a pattern that we hadn't already used
mylm2 <- lm(Y~X4+X2+X3,data=p1Data)#x3 isnt significant in this case
summary(mylm2)

mylm2 <- lm(Y~X4+X2+X2:X4,data=p1Data)#x2 isnt significant in this case
summary(mylm2)

mylm2 <- lm(Y~X4+X2:X4,data=p1Data)#the interaction adds significant change
summary(mylm2)
pairs(cbind(Res=mylm2$residuals,
            fit = mylm2$fitted.values, p1Data),
      panel=panel.smooth, col=as.factor(p1Data$X2))

```
## Part 1 - Recovering a True Model (Beginner Level)

We believe the true model to be...

$$
  Y_i = \underbrace{\beta_0 + \beta_1 X_{4i} + \beta_2 X_{4i} X_{2i}}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = ...-0.7065 +2.3958 X_{4i} + 1.7763X_{4i}X_{2i}
$$

with our estimate of $\sigma$ as 13.11.

```{r}

```


## Part 2 - Recovering a True Model (Intermediate Level)

We believe the true model to be...

$$
  Y_i = \underbrace{...}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = ...
$$

with our estimate of $\sigma$ as ...

```{r}
pairs(p2Data, panel=panel.smooth)

mylm1 <- lm(Y~X5,data=p2Data)
summary(mylm1)

plot(mylm1$residuals~.,p2Data)


mylm2 <- lm(Y~X5 +I(X5^2) + X3,data=p2Data)
summary(mylm2)




```

```{r}
install.packages("plotly")
library(plotly)
#Perform the multiple regression
air_lm <- mylm2 <- lm(Y~X5 +I(X5^2) + X3,data=p2Data)

#Graph Resolution (more important for more complex shapes)
graph_reso <- 0.5

#Setup Axis
axis_x <- seq(min(p2Data$X5), max(p2Data$X5), by = graph_reso)
axis_y <- seq(min(p2Data$X3), max(p2Data$X3), by = graph_reso)

#Sample points
air_surface <- expand.grid(X5 = axis_x, X3 = axis_y, KEEP.OUT.ATTRS=F)
air_surface <- air_surface %>% mutate(Z=predict.lm(air_lm, newdata = air_surface))
air_surface <- acast(air_surface, Month ~ Temp, value.var = "Z") #y ~ x

#Create scatterplot
plot_ly(airquality, 
        x = ~X5, 
        y = ~X3, 
        z = ~Y,
        text = rownames(airquality), 
        type = "scatter3d", 
        mode = "markers") %>%
  add_trace(z = air_surface,
            x = axis_x,
            y = axis_y,
            type = "surface")
```

## Part 3 - Recovering a True Model (Advanced Level)

We believe the true model to be...

$$
  Y_i = \underbrace{...}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = ...
$$

with our estimate of $\sigma$ as ...

```{r}

```
